## Reproducible Research Peer Assessment 1  

========================================================

### Introduction
This markdown document is written to satisfy the requirements for [Peer Assessment 1] for [Reproducible Research] (https://github.com/DataScienceSpecialization/courses/tree/master/05_ReproducibleResearch)  

### Loading and Reprocessing Data
```{r}
activity_raw<-read.csv("activity.csv")
str(activity_raw)
activity<-activity_raw
activity$date<-as.Date(activity$date, format="%Y-%m-%d")  #date fr factor 2 date
activity$steps<-as.numeric(activity$steps)
str(activity)
sum_daily_vector<-aggregate(activity$steps~ activity$date, FUN="sum" )
mean_interval_vector<-aggregate(activity$steps~activity$interval, FUN="mean")
```
### What is the mean total number of steps per day?
```{r}
hist(sum_daily_vector[,2], main="Histogram of steps per day", xlab="Steps")
mean(sum_daily_vector[,2])
median(sum_daily_vector[,2])
```


### What is the daily activity pattern? (This part done w/o inputting missings as the instructions were unclear)
```{r}
plot(mean_interval_vector[,1], mean_interval_vector[,2], type="l", xlab="Interval", ylab="Mean Steps")
record<-mean_interval_vector[mean_interval_vector[,2]==max(mean_interval_vector[,2]),]
print("Interval at max number of steps:")
record[1,1]
```

### Inputating of Missing Values  
1. Calculate number of missing values in data set:
```{r}
steps.summary<-summary(activity$steps)
steps.summary[7]
interval.summary<-summary(activity$interval)
interval.summary[7]
```

2,3 Replace the missing values with the average for the interval and create new dataset
```{r}
activity_mvfi<-activity #copy activity to activity with missing values filled in.
for (i in 1:length(activity_mvfi$date)){
  if (is.na(activity_mvfi$steps[i])){
    record<-mean_interval_vector[mean_interval_vector[,1]== activity_mvfi$interval[i],]
    activity_mvfi$steps[i]<-record[1,2]
    }
}
```
4. Calculate the mean and median with the missing values filled in
```{r}
sum_daily_vector<-aggregate(activity_mvfi$steps~ activity_mvfi$date, FUN="sum")
hist(sum_daily_vector[,2], main="Histogram of steps per day", xlab="Steps")
mean(sum_daily_vector[,2])
median(sum_daily_vector[,2])
```

### Are there differences in activity patterns between weekdays and weekends?
1. Create new vector of variable indicating whether the day was weekday or weekend
```{r}
activity_mvfi$day<-weekdays(activity_mvfi$date)
activity_mvfi$weekdayorend = "weekday"
for (i in 1:length(activity_mvfi$date)){
  if (activity_mvfi$day[i]=="Saturday"){activity_mvfi$weekdayorend[i]="weekend"}
  if (activity_mvfi$day[i]=="Sunday")  {activity_mvfi$weekdayorend[i]="weekend"}
}
```

2. Make panel plot contrasting weekdays and weekends steps taken
```{r}
activity_mvfi_wd<-activity_mvfi[activity_mvfi$weekdayorend=="weekday",]
activity_mvfi_we<-activity_mvfi[activity_mvfi$weekdayorend=="weekend",]
mean_interval_vector_wd<-aggregate(activity_mvfi_wd$steps~activity_mvfi_wd$interval, FUN="mean")
mean_interval_vector_we<-aggregate(activity_mvfi_we$steps~activity_mvfi_we$interval, FUN="mean")
#data_for_lattice<-cbind(mean_interval_vector_wd$interval,mean_interval_vector_wd$steps, mean_interval_vector_we$steps)
plot(mean_interval_vector_we[,1], mean_interval_vector_we[,2], type="l", xlab="Interval", ylab="Mean Steps", main="weekend")
plot(mean_interval_vector_wd[,1], mean_interval_vector_wd[,2], type="l", xlab="Interval", ylab="Mean Steps", main="weekday")
```





